---
title: "Predicting Attitudes Towards Racist Professors"
author: "Georgianna James"
date: '2022-03-02'
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



## Required Packages

```{r packages, message=FALSE}


library(rcfss)
library(tidyverse)
library(tidymodels)
library(rsample)
library(knitr)
library(ranger)
library(kknn)
library(glmnet)

set.seed(100)

theme_set(theme_minimal())
```

## Load the data

```{r load data}
data("gss", package = "rcfss")

# create a data set that omits columns that are not predictors

gss_pred <- gss %>%
  select(-id, -wtss)
```

# Creating training and testing sets

```{r q1 split data}


# split data into train and test

gss_split <- initial_split(gss_pred, prop = 3 / 4)
gss_train <- training(gss_split)
gss_test <- testing(gss_split)
```


# Predictions using a logistic regression model

```{r q2 log model}

# define logistic regression model

log_mod <- logistic_reg() %>%
  set_engine("glm") %>%
  set_mode("classification")
```


```{r q2 cross validation}

# cross validate with 10 folds


# creating folds

gss_fold <- vfold_cv(data = gss_train, v = 10, strata = colrac)
```

```{r q2 fit model}
# refitting model to samples and collecting metrics

gss_fold_pred <- log_mod %>%
  fit_resamples(colrac ~ age + black + degree + partyid_3 + sex + south, resamples = gss_fold) %>%
  collect_metrics()
```


## Logistic Regression Metrics Table

```{r q2 metrics table, echo=FALSE}
kable(gss_fold_pred, col.names = c("Metric", "Estimator", "Mean", "n", "Standard Error", "Config"))
```




# Predictions using a random forest model

```{r q3 define recipe}
# define gss recipe

gss_rec <- recipe(colrac ~ ., data = gss_pred) %>%
  step_impute_median(all_numeric_predictors()) %>%
  step_impute_mode(all_nominal_predictors()) %>%
  step_naomit(all_outcomes(), skip = TRUE)
```

```{r q3define model}

# define random forest model

rf_mod <- rand_forest(
  mode = "classification",
  engine = "ranger",
  mtry = NULL,
  trees = NULL,
  min_n = NULL
)
```

```{r q3 define workflow}

# define workflow

rf_wf <- workflow() %>%
  add_recipe(gss_rec) %>%
  add_model(rf_mod)
```

```{r q3 cross validate}
# cross validate

gss_fold <- vfold_cv(data = gss_train, v = 10, strata = colrac)
```


```{r q3 fit wf }

# fit the workflow

rf_wf_results <- rf_wf %>%
  fit_resamples(resamples = gss_fold) %>%
  collect_metrics()
```

## Random Forrest  Metrics Table

```{r q3 rf results, echo=FALSE}
# present results

kable(rf_wf_results, col.names = c("Metric", "Estimator", "Mean", "n", "Standard Error", "Config"))
```

# Predictions using k-nearest neighbors mdoel

```{r q4 define model}
# define the model

knn_mod <- nearest_neighbor() %>%
  set_engine("kknn") %>%
  set_mode("classification")
```

```{r q4 define recipe}

# define the recipe

knn_rec <- recipe(colrac ~ ., data = gss_pred) %>%
  step_impute_median(all_numeric_predictors()) %>%
  step_impute_mode(all_nominal_predictors()) %>%
  step_naomit(all_outcomes(), skip = TRUE) %>%
  step_novel(all_nominal_predictors()) %>%
  step_dummy(all_nominal_predictors()) %>%
  step_zv(all_predictors()) %>%
  step_normalize(all_numeric())
```

```{r q4 define wf}

# define the workflow

knn_wf <- workflow() %>%
  add_recipe(knn_rec) %>%
  add_model(knn_mod)
```


```{r q4 cross validate}
# cross validate

gss_fold <- vfold_cv(data = gss_train, v = 10, strata = colrac)
```


```{r q4 fit wf}
# fit the workflow

knn_wf_results <- knn_wf %>%
  fit_resamples(resamples = gss_fold) %>%
  collect_metrics()
```


## K-Nearest Neighbors Metrics Table

```{r q4 metrics, echo=FALSE}

kable(knn_wf_results, col.names = c("Metric", "Estimator", "Mean", "n", "Standard Error", "Config"))
```
# Predictions using a Ridge Logistic Regression Model

```{r q5 define model}

# define model

ridge_log_mod <- logistic_reg(
  penalty = .01,
  mixture = 0,
  mode = "classification",
  engine = "glmnet"
)
```

```{r q5 define recipe}

# define recipe

ridge_rec <- recipe(colrac ~ ., data = gss_pred) %>%
  step_impute_median(all_numeric_predictors()) %>%
  step_impute_mode(all_nominal_predictors()) %>%
  step_naomit(all_outcomes(), skip = TRUE) %>%
  step_novel(all_nominal_predictors()) %>%
  step_dummy(all_nominal_predictors()) %>%
  step_zv(all_predictors()) %>%
  step_normalize(all_numeric())
```

```{r q5 define wf}

# define the workflow

ridge_wf <- workflow() %>%
  add_recipe(ridge_rec) %>%
  add_model(ridge_log_mod)
```


```{r q5 cross validate}
# cross validate

gss_fold <- vfold_cv(data = gss_train, v = 10, strata = colrac)
```


```{r q5 fit wf}
# fit the workflow

ridge_wf_results <- ridge_wf %>%
  fit_resamples(resamples = gss_fold) %>%
  collect_metrics()
```


## Ridge Logistic Regression Model Metrics Table

```{r q5 metrics, echo=FALSE}

kable(ridge_wf_results, col.names = c("Metric", "Estimator", "Mean", "n", "Standard Error", "Config"))
```
# Comparing all results

## Logistic Regression Metrics Table

```{r q6 logistic metrics, echo=FALSE}
kable(gss_fold_pred, col.names = c("Metric", "Estimator", "Mean", "n", "Standard Error", "Config"))
```
## Random Forrest  Metrics Table

```{r q6 rf metrics, echo=FALSE}
# present results

kable(rf_wf_results, col.names = c("Metric", "Estimator", "Mean", "n", "Standard Error", "Config"))
```
## 5-Nearest Neighbors Metrics Table

```{r q6  knn metrics, echo=FALSE}

kable(knn_wf_results, col.names = c("Metric", "Estimator", "Mean", "n", "Standard Error", "Config"))
```
## Ridge Logistic Regression Model Metrics Table

```{r q6 ridge metrics, echo=FALSE}

kable(ridge_wf_results, col.names = c("Metric", "Estimator", "Mean", "n", "Standard Error", "Config"))
```



The random forest model had the most accurate predictions, so I will train this recipe/model using the full training set and report the accuracy using the held-out test set of data.

```{r}

# training the rf recipe using the full training set and predicting test set outcomes

rf_test_results <- rf_wf %>%
  fit(data = gss_train) %>%
  predict(new_data = gss_test) %>%
  mutate(true_gss = gss_test$colrac) %>%
  accuracy(truth = true_gss, estimate = .pred_class)
```

## Final Random Forest Model Prediction Results 
```{r echo=FALSE}
kable(rf_test_results, col.names = c("Metric", "Estimator", "Estimate"))
```
